Spring haul-out behavior of seals in the Bering and Chukchi seas

### 2023-09-06 analysis/paper/London_HauloutBeahvior.Rmd
finalize model specification & data
did some exploration of various model specifications (what if day^4, what if less interactions, etc) and still feel this is the best specification

For all seal species, we considered the following 
variables when modeling the hourly haul-out behavior: day-of-year, solar hour,
temperature, wind speed, barometric pressure, precipitation, wind chill
(represented by a *wind:temperature* interaction [@udevitz2009]), and
day-of-year and time-of-day (solar hour) interactions to permit diurnal patterns
to change from March to July. Ribbon and spotted seal models included age-sex
class and interactions between day-of-year and age-sex class, but we omitted
these from bearded seal models due to poor representation of age-sex classes
(Table \@ref(tab:tblDeploy)). Bearded seal models included a latitudinal effect
(and an interaction with day-of-year), since bearded seals occupy a substantial
range during the spring and we were interested in possible differences in the
timing of haul-out behavior along a latitudinal gradient. We omitted the
latitudinal effect from ribbon and spotted seal models because, during the
spring, these species are most prevalent near the southern ice edge in the
Bering Sea [@conn2014a].

@@

We followed Ver Hoef et al. [-@verhoef2014a] in using linear, quadratic, and
cubic effects of day-of-year to represent temporal changes in behavior. However,
unlike previous models for harbor seals [@london2012] and ice-associated seals
[@verhoef2014a], which treated hour-of-day as a 24-level categorical variable to
capture diurnal cycles, we adopted a continuous formulation based on Fourier
series that provides a flexible model while preserving the inherent circularity
needed for time-of-day effects (i.e., hour 0 should be equal to hour 24). It
also represents hour-of-day with 6 parameters, which is a considerable reduction
when compared to a 24-parameter variable, especially when fitting models including
interactions between hour-of-day and other variables (e.g., age-sex class,
day-of-year). According to this approach, we used the following specification
for hour-of-day effects:

@@

options(
  gargle_oauth_cache = here::here(".secrets"),
  gargle_oauth_email = TRUE
)

>>>>

For all seal species, we considered the following variables when modeling the
hourly haul-out behavior: day-of-year, solar hour, temperature, wind speed,
barometric pressure, precipitation, and wind chill (represented by a
*wind:temperature* interaction [@udevitz2009]). Ribbon and spotted seal models
included age-sex class and interactions between day-of-year and age-sex class,
but we omitted these from bearded seal models due to poor representation of
age-sex classes (Table \@ref(tab:tblDeploy)). Bearded seal models included a
latitudinal effect (and an interaction with day-of-year), since bearded seals
occupy a substantial range during the spring and we were interested in possible
differences in the timing of haul-out behavior along a latitudinal gradient. We
omitted the latitudinal effect from ribbon and spotted seal models because,
during the spring, these species are most prevalent near the southern ice edge
in the Bering Sea [@conn2014a].

@@

We followed Ver Hoef et al. [-@verhoef2014a] in using linear, quadratic, and cubic
effects of day-of-year to represent temporal changes
in behavior during the season. However, unlike previous models for harbor seals
[@london2012] and ice-associated seals [@verhoef2014a], which treated
hour-of-day as a 24-level categorical variable to capture diurnal cycles, we
adopted a continuous formulation based on Fourier series that provides a
flexible model while preserving the inherent circularity needed for time-of-day
effects (i.e., hour 0 should be equal to hour 24). It also represents
hour-of-day with 6 parameters, which is a considerable reduction when compared
to a 24-parameter variable. According to this approach, we used the following
specification for hour-of-day effects:

@@

options(
  gargle_oauth_cache = here::here(".secrets"),
  gargle_oauth_email = "josh.london@noaa.gov"
)
